---
title: '[Learning Apache Spark](https://github.com/MingChen0919/learning-apache-spark)'
output:
    html_document:
      highlight: pygments
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_knit$set(progress = FALSE)
knitr::opts_chunk$set(error = TRUE, echo = FALSE)
library(htmltools)
```

```{r, echo=FALSE}
# to make the css theme to work, <link></link> tags cannot be added directly 
# as <script></script> tags as below.
# it has to be added using a code chunk with the htmltool functions!!!
css_link = tags$link()
css_link$attribs = list(rel="stylesheet", href="vakata-jstree-3.3.5/dist/themes/default/style.min.css")
css_link
```

```{r, eval=FALSE, echo=FALSE}
# this code chunk is purely for adding comments
# below is to add jQuery and jstree javascripts
```
<script src="https://code.jquery.com/jquery-3.3.1.min.js"></script>
<script src="vakata-jstree-3.3.5/dist/jstree.min.js"></script>

```{r, eval=FALSE, echo=FALSE}
# this code chunk is purely for adding comments
# javascript code below is to build the file tree interface
# see this for how to implement opening hyperlink: https://stackoverflow.com/questions/18611317/how-to-get-i-get-leaf-nodes-in-jstree-to-open-their-hyperlink-when-clicked-when
```
<script>
  $(function () {
    // create an instance when the DOM is ready
    $('#jstree').jstree().bind("select_node.jstree", function (e, data) {
     window.open( data.node.a_attr.href, data.node.a_attr.target )
    });
  });
</script>


```{r}
file_tree = function(dir = '.'){
  # # get the OUTPUT_DIR folder data: dataset_NUMBER_files
  # report_files_path = Sys.getenv('REPORT_FILES_PATH')
  # output_dir = tail(strsplit(report_files_path, '/')[[1]], 1)
  
  files = list.files(path = dir, recursive = FALSE, full.names = TRUE)
  # files also include directorys, need to remove directorys
  files = files[!dir.exists(files)]
  dirs = list.dirs(path = dir, recursive = FALSE, full.names = TRUE)
  # exclude .ipynb_checkpoints folder
  # ipynb_checkpoints = grep(pattern = 'ipynb_checkpoints', x = dirs)
  # dirs = dirs[-ipynb_checkpoints]
  github_repo_url = 'https://github.com/MingChen0919/learning-apache-spark/blob/master/'
  tags$ul(
    {
      if (length(files) > 0) {
        lapply(files, function(x){
          path_end = tail(strsplit(x, '/')[[1]],1)
          li_item = tags$li(tags$a(path_end, href=paste0(github_repo_url, x)))
          li_item$attribs = list('data-jstree'='{"icon":"jstree-file"}')
          li_item
        })
      }
    },
    {
      if (length(dirs) > 0) {
        lapply(dirs, function(x){
          path_end = tail(strsplit(x, '/')[[1]],1)
          if (!(path_end %in% c('vakata-jstree-3.3.5', '.ipynb_checkpoints', 'spark-warehouse', 'ipynb'))) {
            li_item = tags$li(path_end, file_tree(x))
            li_item$attribs = list('data-jstree' = '{"icon":"jstree-folder"}', class=list('jstree-open'))
            li_item
          }
        })
      }
    }
  )
}
```


**[Ming Chen](https://github.com/MingChen0919) & [Wenqiang Feng](http://web.utk.edu/~wfeng1/)**

## Introduction

This repository contains mainly notes from learning Apache Spark by [Ming Chen](https://github.com/MingChen0919) & [Wenqiang Feng](http://web.utk.edu/~wfeng1/). We try to use the detailed demo code and examples to show how to use pyspark for big data mining. **If you find your work wasn't cited in this note, please feel free to let us know.**


## Cheat Sheets

+ [Python Basics](https://s3.amazonaws.com/assets.datacamp.com/blog_assets/PythonForDataScience.pdf)
+ [Pandas Basics](http://datacamp-community.s3.amazonaws.com/3857975e-e12f-406a-b3e8-7d627217e952)
+ [Numpy Basics](https://s3.amazonaws.com/assets.datacamp.com/blog_assets/Numpy_Python_Cheat_Sheet.pdf)
+ [Scikit-Learn](http://datacamp-community.s3.amazonaws.com/5433fa18-9f43-44cc-b228-74672efcd116)
+ [RDD Basics](https://s3.amazonaws.com/assets.datacamp.com/blog_assets/PySpark_Cheat_Sheet_Python.pdf)
+ [DataFrame Basics](https://s3.amazonaws.com/assets.datacamp.com/blog_assets/PySpark_SQL_Cheat_Sheet_Python.pdf)
+ [Apache Spark Cheat Sheet](https://hackr.io/tutorials/learn-apache-spark)
    
## Contents

```{r, echo=FALSE}
# create a div container to store the file tree interface
tags$div(
  id="jstree",
  file_tree('notebooks')
)
```


## Feedback and suggestions

Your comments and suggestions are highly appreciated. We are more than happy to receive corrections, suggestions or feedbacks for improvements.
